---
title: "All figures for normalization manuscript"
author: "Hanneke Leegwater"
date: "`r Sys.Date()`"
output:   
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = stringr::str_replace(inputFile, ".Rmd",paste0("_", stringr::str_remove_all(Sys.Date(),"-"),'.html'))) })
---

This script reproduces all figures in the manuscript. The data directory needs to point to the correct location where the Calculations script has created files.

# Preparations

## Load R libraries

```{r}
library(tidyverse)
library(here)
library(cowplot)
library(pheatmap)
library(ggvenn)
library(svglite)
library(dendsort)
source(here("code/functions/data_analysis_plots.R"))
```

## Define input and output directory

```{r}
data_dir <- here("manuscript/data_for_figures")

output_dir <- here("manuscript", str_glue("figures_v{str_remove_all(Sys.Date(), '-')}"))
print(output_dir)
if(!dir.exists(output_dir)) dir.create(output_dir)
```

## Load all metadata

```{r}
# Sample metadata
metadata <- read_csv(here(data_dir, "metadata.csv"))
 
print(dim(metadata))
print(head(metadata))

### Information on lipid classes
lipid_info <- read_csv(here(data_dir, "lipid_info.csv"))

print(head(lipid_info))
print(dim(lipid_info))
```

## Define all colors used in the manuscript

```{r}
color_legend_morphology <- list(
  "Epithelial-like" = "#FDE725", # scales::viridis_pal()(5)[5],
  "Mesenchymal-like" = "#3B528B", # scales::viridis_pal()(5)[1],
  "Round" = "#CB181D",
  "Suspension" = "#5DC863",
  "Unknown" = "grey",
  "E" = "#FDE725", 
  "M" = "#3B528B", 
  "R" = "#CB181D",
  "S" = "#5DC863",
  "U" = "grey"
)

color_legend_icc <- c(
  "data-based" = "#0099FF", 
  "biology-based" = "#FF9B2D", 
  "not normalized" = "black",
  "Not" = "black",
  "Prot conc." = "#FF9B2D",
  "Cell count" = "#F35C19",
  "Sum" = "#0099FF",
  "Median" = "#A5B2E3",
  "PQN" = "#001158"
)

color_legend_measurement_modes <- list(mode = c(
  "Positive" = "#F46E31",
  "Negative" = "#001158",
  "Switching" = "#A5B2E3"
))

color_legend_triglycerides <- list(
  "triglyceride" = "#001158",
  "other lipid class" = "#F46E31"
)
```

## Define order to use when showing normalizations

```{r}
normalization_order <- c("Not", "Prot conc.", "Cell count", "Sum", "Median", "PQN")
```

## Define personal theme to use for all plots

Adapted from tutorial: Harvard Chan Bioinformatics Core (HBC) with CC BY 4.0, <https://hbctraining.github.io/publication_perfect/>

and <https://ggplot2.tidyverse.org/articles/faq-customising.html#what-is-the-default-size-of-geom_text-and-how-can-i-change-the-font-size-of-geom_text>

```{r}
update_geom_defaults("label", list(size = 3, size.unit = "pt"))
# Most axis labels are 8, ticks on the axis are 6
fontsize_smaller <- 6
fontsize_larger <- 8
# Font size of all figure labels is 10
figure_label_size <- 10
personal_theme <- function(){ 
  theme_half_open(font_size = 10) +
  theme(legend.position = "none") +
  # White background
   theme(plot.background = element_rect(fill = "white", color = NA),
         axis.title.x = element_text(size = fontsize_larger),
         axis.text.x = element_text(size = fontsize_smaller),
         axis.title.y = element_text(size = fontsize_larger),
         axis.text.y = element_text(size = fontsize_smaller),
         legend.text  = element_text(size=fontsize_smaller), 
         legend.title = element_text(size=fontsize_larger),
         plot.title = element_text(size=fontsize_larger)
          )
}

plot_grid <- purrr::partial(cowplot::plot_grid, label_size = figure_label_size)

# Slightly smaller dot sizes in violin/box plots
dot_scaling_size <- 0.7
update_geom_defaults("point", list(size = dot_scaling_size))
geom_boxplot <- purrr::partial(ggplot2::geom_boxplot, outlier.size = dot_scaling_size)
```

## Save plots as png, svg, eps, tiff and pdf

Using the cowplot save_plot function but with multiple file types.

Journal of chemometrics says: 
-   Black and white and colour photos - 300 dpi
-   Graphs, drawings, etc - 800 dpi preferred; 600 dpi minimum
-   Combinations of photos and drawings (black and white and colour) - 500 dpi
-   Please supply artwork at the intended size for printing, the maximum reproduction size of each illustration is 140 mm × 200 mm.

```{r}
page_width <- 14 #14 cm instead of default in inch.
height_max <- 20
save_as_png_svg_pdf <- function(figure_name, plot_object, base_height = 8, base_width=page_width){
  if(base_height > height_max) warning("Base height is more than max page height")
  for(file_extension in c(".svg", ".png", ".pdf", ".eps")){#, ".tiff")){
    save_plot(
      here(output_dir, str_c(figure_name, file_extension)),
      plot_object,
      base_height = base_height/cm(1),
      base_width = base_width/cm(1), dpi = 800)    
  }
}
```

# Main figures

## Figure 1

Figure 1. Variation in data is mostly influenced by median signal in a sample.

A PCA was calculated using log2 transformed, centered, and scaled data. (A) Morphology was not the basis for the observed separation in the PCA; epithelial-like (yellow), mesenchymal-like (blue), round (red), suspension (green), and unknown morphology (grey). (B) Example cell lines highlighted to show that replicates do not cluster. (C) Median detected signal for the sample drives separation in PC1, which explained 59% of total variation. (D) PC1 and log2 median signal per sample showed a strong correlation negative to each other (rs=-0.99).

### Load data

```{r}
load(here(data_dir, "pca_raw_data.RData"))
medians_per_sample <- read_csv(here(data_dir, "medians_per_sample.csv"))
```

### Write functions

```{r}
plot_PCA_df <- function(pca_obj, metadata, color_metadata, legend = T,
                        components = c("PC1","PC2")){
  PC1_var <- round(pca_obj$prop_expl_var$X[[components[1]]], 2) *100
  x_label <- str_glue("{components[1]} ({PC1_var}%)")
  PC2_var <- round(pca_obj$prop_expl_var$X[[components[2]]],2) *100
  y_label <- str_glue("{components[2]} ({PC2_var}%)")
  df <- get_PCA_df(pca_obj) %>%
    left_join(metadata, by = "Original_name") %>%
    arrange(Morphology)
  
  p <- ggplot(df, aes(x = .data[[components[1]]], y = .data[[components[2]]],
                      color = .data[[color_metadata]])) +
    personal_theme() +
    labs(x = x_label, y = y_label) +
    theme(legend.position = ifelse(legend, "right", "none"))
  
  if(color_metadata != "Morphology"){
    p <- p + geom_point()
    } else{
    p <- p + geom_point() + 
      scale_color_manual(values = color_legend_morphology)
  }
  return(p)
}

plot_PCA_replicate_groups <- function(pca_obj, metadata, to_highlight_legend, shape_legend = NULL,
                        legend = FALSE, components = c("PC1","PC2"), title = NULL){
  
  # Turn mixomics pca object into data frame for ggplot
  PCx_var <- round(pca_obj$prop_expl_var$X[[components[1]]], 2) *100
  x_label <- str_glue("{components[1]} ({PCx_var}%)")
  PCy_var <- round(pca_obj$prop_expl_var$X[[components[2]]],2) *100
  y_label <- str_glue("{components[2]} ({PCy_var}%)")
  df_pca <- get_PCA_df(pca_obj)

  
  # Add metadata
  df_pca <- left_join(df_pca, metadata, by = "Original_name") %>%
    mutate(`Cell line` = ifelse(group %in% names(to_highlight_legend), group, "Other")) %>% 
    mutate(`Cell line` = factor(`Cell line`)) %>%
    mutate(`Cell line` = relevel(`Cell line`, "Other")) %>%
    arrange(`Cell line`)
  
  # Not highlighted part
  p <- ggplot(filter(df_pca, `Cell line` == "Other"), 
              aes(x = .data[[components[1]]], y = .data[[components[2]]])) +
    geom_point(aes(fill = `Cell line`, color = `Cell line`)) +
    geom_line(aes(group = group, color = `Cell line`))
  # Add highlighted part
  if(!is.null(shape_legend)){
  p <- p + 
    geom_point(data = filter(df_pca, `Cell line` != "Other"),
               aes(fill = `Cell line`, color = `Cell line`, shape = `Cell line`))
  } else{
  p <- p + 
    geom_point(data = filter(df_pca, `Cell line` != "Other"),
               aes(fill = `Cell line`, color = `Cell line`), shape = 15) # squares for highlighted cell lines.
  }
  p <- p +
    personal_theme() +
    theme(legend.position = ifelse(legend, "bottom", "none")) +
    labs(x = x_label, y = y_label) +
    geom_line(data = filter(df_pca, `Cell line` != "Other"), aes(group = group, color = `Cell line`)) +
    scale_color_manual(values = to_highlight_legend)
  if(!is.null(shape_legend)) p <- p + scale_shape_manual(values = shape_legend)
  
  if(!is.null(title)) p <- p + ggtitle(title)
  return(p)
}
```

```{r}
plot_PCA_median <- function(pca_obj, metadata, color_metadata, legend = T){
  # Turn mixomics pca object into data frame for ggplot
  PC1_var <- round(pca_obj$prop_expl_var$X[["PC1"]], 2) *100
  x_label <- str_glue("PC1 ({PC1_var}%)")
  PC2_var <- round(pca_obj$prop_expl_var$X[["PC2"]],2) *100
  y_label <- str_glue("PC2 ({PC2_var}%)")
  
  df <- get_PCA_df(pca_obj) %>%
    left_join(metadata, by = "Original_name")
  print(str_glue("Correlation between PC1 and median signal is: {cor(df[[color_metadata]], df$PC1)}"))
  print(str_glue("Correlation between PC2 and median signal is: {cor(df[[color_metadata]], df$PC2)}"))
  p <- ggplot(df, aes(x = PC1, y = PC2, 
                      color = .data[[color_metadata]])) +
    geom_point() +
    personal_theme() +
    scale_color_gradient(name = "log2 median", low = "#D9D9D9", high = "black") +
    theme(legend.position = ifelse(legend, "right", "none")) +
    labs(x = x_label, y = y_label, legend = "log2 median") 
  return(p)
}
```

### Plot log2 transformed points

```{r fig.height=2.6, fig.width=4}
fig1a <- plot_PCA_df(pca_raw_data, metadata, "Morphology") +
   scale_color_manual(values = color_legend_morphology)
fig1a + guides(color = guide_legend(override.aes = list(size = 5, shape = 15)))

the_legend_1a <- get_plot_component(
  plot_PCA_df(pca_raw_data, metadata, "Morphology") + 
    guides(color = guide_legend(), shape = FALSE) +
    theme(legend.position = "right"),
  "guide-box-right"
)
```

### Plot replicates

```{r fig.height=2.6, fig.width=4}
to_highlight_legend = c(
    "MCF-7" = "#3B528BFF",
    "MDA-MB-175VII" = "#CB181D",
    "HCC38" = "#E97132", 
    "MDA-MB-436" = "#196B24",
    "OCUB-F" = "#A02B93",
    "Other" = "grey"
  )
fig1b <- plot_PCA_replicate_groups(pca_obj = pca_raw_data, metadata, to_highlight_legend = to_highlight_legend)

fig1b
fig1b + theme(legend.position = "right") #+ guides(color = guide_legend(override.aes = list(size = 5, shape = 15)))

the_legend_1b <- get_plot_component(
  fig1b + 
    guides(color = guide_legend(),
           shape = FALSE) +
    theme(legend.position = "right"),
  "guide-box-right"
)
```

### Color by median signal

```{r fig.height=2.8, fig.width=5.5}
median_signal_per_sample <- dplyr::select(medians_per_sample, Original_name, log2median)
fig1c <- plot_PCA_median(pca_raw_data, median_signal_per_sample, "log2median")
fig1c + theme(legend.position = "right")
```

Correlate PC1 to median signal. `plot_median_and_PC1` creates a plot with some median value and the PC1, to show 
whether the first principal component is biased by the original median of a sample.

```{r}
plot_median_and_PC1 <- function(pca_obj, normalization_name, median_df){
  PC1_var <- round(pca_obj$prop_expl_var$X[["PC1"]],2) *100
  x_label <- str_glue("PC1 ({PC1_var}%)")
  
  df <- pca_obj$variates$X %>% data.frame() %>%
    rownames_to_column("Original_name") %>%
    left_join(median_df, by = "Original_name") 
  
  cor_nr <- round(cor(df$PC1, df$log2median, method = "spearman"), 2)
  print(str_c("Spearman correlation between PC1 and log2 median signal for ", normalization_name, " is: ",cor_nr))
  
  p <- df %>%
    ggplot(aes(x = PC1, log2median)) + 
    geom_point(color = "grey20") +
    personal_theme() +
    labs(y = "log2 median abundance", x = x_label) +
    geom_label(x = Inf, y = Inf, hjust = 1, vjust = 1,
              label = str_glue("r[s] == {cor_nr}"),
              label.size = NA, parse = T)
  return(p)
}
fig1d <- plot_median_and_PC1(pca_raw_data, "raw data", median_signal_per_sample)
fig1d
```

### Write to file

```{r fig.height=3.2, fig.width=5.5}
all_plots_1 <- plot_grid(
  plotlist = align_plots(
    fig1a + theme(legend.position = "right"), 
    fig1b + theme(legend.position = "right"), 
    fig1c + theme(legend.position = "right"), 
    fig1d + theme(legend.position = "none"),
    align="hv", axis="tblr"
    ),
    labels = LETTERS[1:4],
    nrow = 2)
all_plots_1

save_as_png_svg_pdf(
  "Figure1",
  all_plots_1,
  base_height = 9)

save_plot(
  here(output_dir, "Figure1a_legend.png"),
  the_legend_1a)

save_plot(
  here(output_dir, "Figure1b_legend.png"),
  the_legend_1b)
```

## Figure 2

Figure 2. Lipid abundance distributions are not directly related to total protein concentration or cell count.

(A) Distribution of lipid abundances for 5 cell lines where biological replicates have the most similar total protein concentration with a standard deviation (sd) 0.00 -0.01 µg/µL, or (B) most different total protein concentrations with a sd 0.15 - 0.30 µg/µL. Distributions for cell counts with (C) minimum sd 0.11 - 0.18 million cells or (D) maximum sd for 1.7 to 2.1 million cells. Distributions are colored per cell morphology with epithelial-like (yellow), mesenchymal-like (blue), round (red), or suspension (green) morphology. Correlation between the two biology-based methods (E) or between cell count (F) or protein concentration (G) and log2 median sample abundance. rs: Spearman correlation coefficient

### Load data

```{r}
sd_biological_measurements <- read_csv(here(data_dir, "sd_biological_measurements.csv"))
raw_data_log2 <- read_csv(here(data_dir, "raw_data_log2.csv"))
```

### Show 5 cell lines with similar and different median signals.

Plots partially based on <http://www.sthda.com/english/wiki/ggplot2-violin-plot-quick-start-guide-r-software-and-data-visualization>.

```{r}
plot_five_sample_distributions <- function(df, for_filter, plot_title = NULL, max_variation = T, color_legend = NULL, free_y = F){
  
  df <- df %>%
    pivot_longer(!Original_name, names_to = "lipid", values_to = "value") %>%
    left_join(metadata, by = "Original_name") %>%
    filter(group %in% for_filter) %>%
    # sort by morphology
    mutate(Morphology = factor(Morphology, levels = names(color_legend_morphology)))
  to_sort <- df %>% 
    dplyr::select(group, Morphology) %>%
    arrange(Morphology) %>% unique()
  df <- df %>%
    mutate(group = factor(group, levels = to_sort$group),
           Original_name = factor(Original_name, levels = unique(Original_name)))
  
  if(is.null(color_legend)){
    p <- df %>%
      ggplot(aes(group = Original_name, x = group, y = value, fill = Morphology)) +
        scale_fill_manual(values = color_legend_morphology)
  } else{
    p <- df %>%
      ggplot(aes(group = Original_name, x = group, y = value, fill = group)) +
        scale_fill_manual(values = color_legend)
  }
  p <- p +
    geom_violin() +
    geom_boxplot(position=position_dodge(0.9), width=0.2, fill = "white") +
    personal_theme() +
    scale_x_discrete(guide = guide_axis(angle = 20)) +
    labs(x = NULL, y = "log2 lipid abundance")

  if(!is.null(plot_title)){
    p <- p + ggtitle(plot_title)
  }
  
  if(!free_y){
    p <- p + scale_y_continuous(limits = c(-15,7))
  }
  return(p)
}
```

```{r fig.height=3, fig.width=5.5}
max_sd <- sd_biological_measurements %>%
  slice_max(sd_protein_concentration_mg_per_mL, n = 5)
max_sd

min_sd <- sd_biological_measurements %>%
  slice_min(sd_protein_concentration_mg_per_mL, n = 5)
min_sd
fig2a <- raw_data_log2 %>%
  plot_five_sample_distributions(for_filter = min_sd$group, max_variation = F)
  
fig2b <- raw_data_log2 %>%
  plot_five_sample_distributions(for_filter = max_sd$group, max_variation = T)

the_legend_2ab <- get_plot_component(
  plot_PCA_df(pca_raw_data, metadata, "Morphology") + 
    theme(legend.position = "bottom"), 
  "guide-box-bottom"
)

all_plots_2ab <- plot_grid(fig2a, fig2b,
  align = 'hv',
  labels = c("A", "B"),
  hjust = -1,
  nrow = 1
)
all_plots_2ab
```

Repeat for cell count

```{r fig.height=2.6, fig.width=5.5}
max_sd <- sd_biological_measurements %>%
  slice_max(sd_cell_nr_10E6, n = 5)

min_sd <- sd_biological_measurements %>%
  slice_min(sd_cell_nr_10E6, n = 5)

fig2c <- raw_data_log2 %>%
  plot_five_sample_distributions(for_filter = min_sd$group, max_variation = F)
  
fig2d <- raw_data_log2 %>%
  plot_five_sample_distributions(for_filter = max_sd$group, max_variation = T)

all_plots_2cd <- plot_grid(fig2c, fig2d,
  align = 'hv',
  labels = c("C", "D"),
  hjust = -1,
  nrow = 1
)

all_plots_2cd
```

### Add correlation plots

Following advice on the labels on <https://stackoverflow.com/questions/60189086/how-to-make-superscripts-inside-glue>

```{r fig.height=8/3, fig.width=5.5}
plot_correlation <- function(df, x, y){
  df <- df %>%
    rename_with(~str_replace_all(.x, "_", " "))
  cor_nr <- round(cor(df[[x]], df[[y]], method = "spearman"),2)
  print(cor_nr)
  
  p <- df %>%
    ggplot(aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(color ="grey20") +
    personal_theme() +
    geom_smooth(method = "lm", se = FALSE, 
                color = "black", linetype = "dashed", linewidth = dot_scaling_size) +
    geom_label(x = Inf, y = Inf, hjust = 1, vjust = 1,
              label = str_glue("r[s] == {cor_nr}"),
              label.size = NA, parse = T)
  return(p)
}

medians_per_sample <- rename(medians_per_sample,
                             "median abundance" = "median",
                             "log2 median abundance" = "log2median")

all_plots_2efg <- plot_grid(
  plot_correlation(medians_per_sample,
                   x = "cell count (million)", 
                   y = "prot conc. (mg/mL)"),
  plot_correlation(medians_per_sample,
                   x = "cell count (million)", 
                   y = "log2 median abundance"),
  plot_correlation(medians_per_sample,
                   x = "prot conc. (mg/mL)", 
                   y = "log2 median abundance"),
  labels = c("E", "F", "G"),
  rel_widths = c(1,1,1),
nrow = 1
)
all_plots_2efg
```

```{r}
plot_grid(the_legend_2ab)
```

```{r fig.height=8, fig.width=5.5}
all_plots_2abcdefg <- plot_grid(
    all_plots_2ab, all_plots_2cd, 
    plot_grid(NULL, the_legend_2ab, NULL, rel_widths = c(0.2,1,0.2)), 
    all_plots_2efg,
    ncol = 1,
    rel_heights = c(1,1,0.1,1)
  )
all_plots_2abcdefg
```

### Write to file

```{r fig.height=6, fig.width=5.5}
save_as_png_svg_pdf(
  "Figure2",
  all_plots_2abcdefg + theme(plot.background = element_rect(fill = "white", color = "white")),
  base_height = 14)

save_plot(
  here(output_dir, "Figure2_legend.png"),
  the_legend_2ab)
```

## Figure 3

Figure 3. Data-based normalizations improved similarity between biological replicates.

Distributions of intraclass correlations per lipid are calculated before and after normalization for (A) the entire panel of cell lines. Lipid density distributions per sample (B) before normalization, (C) after normalization by protein concentration or (D) after PQN normalization for the cell line panel. (E) Distributions of intraclass correlations per lipid are calculated before and after normalization for cell lines with similar morphologies.

### Load data

And sort by normalization_order

```{r}
ICC_cell_lines_normalization_merged <- read_csv(here(data_dir, "ICC_cell_lines_normalization_merged.csv")) %>%
  rename_with(~str_remove(.x, "_log2Transformed_notScaled")) %>%
  select(lipid, all_of(normalization_order))
head(ICC_cell_lines_normalization_merged)

setdiff(colnames(ICC_cell_lines_normalization_merged), normalization_order)

ICC_treatment_2DG_normalization_merged <- read_csv(here(data_dir, "ICC_treatment_2DG_normalization_merged.csv")) %>%
  rename_with(~str_remove(.x, "_log2Transformed_notScaled")) %>%
  select(lipid, all_of(normalization_order))
head(ICC_treatment_2DG_normalization_merged)

```

```{r}
data_cell_lines_autoScaled <- list()
data_cell_lines_notScaled <- list()

# Loop over all settings and all data frames
for(normalization_mode in normalization_order){
  data_subset <- "cell_lines"
  transformation <- "log2Transformed"
  
  # Define normalization_mode filename
  normalization_name <- switch(
    normalization_mode,
    "Not" = "NotNormalized",
    "Prot conc." = "BCANormalized",
    "Cell count" = "CellCountNormalized",
    "Sum" = "SumNormalized",
    "Median" = "MedianNormalized",
    "PQN" = "PQNNormalized")
  print(str_glue("Reading {data_subset}, {normalization_name}, {transformation}"))
  
  # autoScaled dataset
  file_name <- here(data_dir, "data_normalized", "cell_lines", str_glue("{data_subset}_merged_{normalization_name}_{transformation}_autoScaled.csv"))
  if(!file.exists(file_name)) stop(str_glue("file not found: {file_name}"))
  data_cell_lines_autoScaled[[normalization_mode]] <- read_csv(file_name, show_col_types = F)
  
  # notScaled dataset
  file_name <- here(data_dir, "data_normalized", "cell_lines", str_glue("{data_subset}_merged_{normalization_name}_{transformation}_notScaled.csv"))
  if(!file.exists(file_name)) stop(str_glue("file not found: {file_name}"))
  data_cell_lines_notScaled[[normalization_mode]] <- read_csv(file_name, show_col_types = F)
}
```

### Define plots

Reuse Figure 1B and 1D. Font size based on <https://stackoverflow.com/questions/13297995/changing-font-size-and-direction-of-axes-text-in-ggplot2>

```{r}
# TODO move to calculations script
calculate_PCA <- function(df_name){
  df <- column_to_rownames(data_cell_lines_autoScaled[[df_name]], "Original_name")
  pca_obj <- mixOmics::pca(df, ncomp = ifelse(ncol(df) >=8, 8, ncol(df)))
  return(pca_obj)
}

plot_two_ICC <- function(x, y, df = ICC_cell_lines_normalization_merged){
  p <- df %>%
    ggplot(aes(x = .data[[x]], y = .data[[y]], text = lipid)) + 
      geom_point(alpha = 0.4, stroke = NA) +
      personal_theme() +
      geom_abline(slope = 1, intercept = 0, color = "red") +
      scale_x_continuous(limits = c(0,1)) +
      scale_y_continuous(limits = c(0,1)) +
      labs(x = str_glue("ICC {x} normalized"), y = str_glue("ICC {y} normalized"))
  return(p)
}

make_sample_densityplot <- function(df, title){
  p <- df %>%
    pivot_longer(!Original_name, names_to = "lipid") %>%
    ggplot(aes(x = value, group = Original_name)) +
    geom_density(color = "black") +
    personal_theme() +
    labs(x = "log2 lipid abundance")
  return(p)
}
```

### Old figure 2a

```{r fig.height=2.6, fig.width=2.6}
plot_ICC <- function(all_ICCs){
  to_sort_by <- normalization_order

  all_ICCs %>%
    pivot_longer(!lipid, names_to = "normalization", values_to = "ICC") %>%
    
    # Sort based on the normalization specified above
    mutate(strategy = ifelse(str_detect(normalization, "Median|Sum|PQN"), "data-based",
                               ifelse(str_detect(normalization, "Prot|Cell"), "biology-based", "not normalized")),
           normalization = factor(normalization, levels = to_sort_by)) %>%
    arrange(normalization) -> all_ICCs
    
    # Build the plot
    p1 <-  all_ICCs %>% ggplot(aes(x = normalization, y = ICC, fill = strategy)) + 
      geom_violin() +
      geom_boxplot(position=position_dodge(1), width=0.2, fill = "white") +
      personal_theme() +
      labs(legend = "normalization type") +
      scale_x_discrete(guide = guide_axis(angle = 30)) +
      scale_y_continuous(limits = c(0, 1)) +
      scale_fill_manual(values = color_legend_icc) +
      labs(x = NULL)
  p2 <- all_ICCs %>% ggplot(aes(x = ICC, color = normalization)) + 
      geom_density() +
      personal_theme() +
      labs(legend = "normalization type") +
      scale_x_continuous(limits = c(0, 1)) +
      scale_color_manual(values = c(color_legend_icc[normalization_order]))
  return(list(p1, p2))
}
icc_overlaps <- plot_ICC(all_ICCs = ICC_cell_lines_normalization_merged)
icc_overlaps
```

### Build the plot

```{r fig.height=2, fig.width=12}
## PCA
to_highlight_legend # Same as 1B

all_PCA <- lapply(normalization_order, calculate_PCA)
names(all_PCA) <- normalization_order

plot_list_PCA <- lapply(all_PCA, plot_PCA_replicate_groups, metadata = metadata, 
                    to_highlight_legend = to_highlight_legend)
plot_grid(plotlist = plot_list_PCA, nrow = 1)

### Correlation with median of not-normalized not-scaled sample
plot_list_mediancor <- map2(all_PCA, names(all_PCA), ~plot_median_and_PC1(.x, .y, median_signal_per_sample))
plot_grid(plotlist = plot_list_mediancor, nrow = 1)

### Overlapping densities
plot_list_densities <- map2(lapply(data_cell_lines_notScaled, head, 50), names(data_cell_lines_notScaled), ~make_sample_densityplot(.x, title = .y))
plot_grid(plotlist = plot_list_densities, nrow = 1)
### ICC
plot_list_ICC <- lapply(normalization_order, function(name) plot_two_ICC("Not", name) + ggtitle(name) + theme(plot.title = element_text(color = color_legend_icc[name])))
plot_list_ICC[1] <- icc_overlaps[2]
plot_grid(plotlist = plot_list_ICC, nrow = 1)

### 5 cell line example

plot_list_violins <- map2(data_cell_lines_notScaled, names(data_cell_lines_notScaled), ~plot_five_sample_distributions(.x, for_filter = setdiff(names(to_highlight_legend), "Other"), color_legend = to_highlight_legend, free_y = T) + ggtitle(.y))
plot_grid(plotlist = plot_list_violins, nrow = 1)
```

#### Show full plot

Create with double the size and letters and scale later in pdf.
```{r}
cowplot::plot_grid(plotlist = plot_list_ICC, 
                     nrow = 1, labels = LETTERS[1:2], label_size = 2*figure_label_size,
                     vjust = 1, hjust = 0)
  
```


```{r fig.height=5.5*2, fig.width=5.5*2}
### Show all together.
all_plots_3_largeviolin <- cowplot::plot_grid(
  # A-B
  cowplot::plot_grid(plotlist = plot_list_ICC, 
                     nrow = 1, labels = LETTERS[1:2], label_size = 1.5*figure_label_size,
                     vjust = 1, hjust = 0), 
  # C-E
  plot_grid(plotlist = plot_list_PCA, nrow = 1), 
  plot_grid(plotlist = plot_list_mediancor, nrow = 1), 
  plot_grid(plotlist = plot_list_densities, nrow = 1), 
  # F
  cowplot::plot_grid(plotlist = plot_list_violins[1:3], 
                     nrow = 1, labels = LETTERS[6], label_size = 1.5*figure_label_size, 
                     vjust = 1, hjust = 0),
  plot_grid(plotlist = plot_list_violins[4:6], nrow = 1),
  
  labels = c("", LETTERS[3:5], NULL),
  rel_heights = c(1.1,1,1,1,1.1,1.1),
  label_size = 1.5*figure_label_size,
  vjust = 1, hjust = 0,
  nrow = 6)
all_plots_3_largeviolin
```

### Calculate nr of lipids with high ICC

Show \> 0.75

```{r}
ICC_treatment_2DG_normalization_merged %>%
  column_to_rownames("lipid") %>%
  mutate(across(everything(), function(x) x > 0.75)) %>%
  colSums()

ICC_cell_lines_normalization_merged %>%
  column_to_rownames("lipid") %>%
  mutate(across(everything(), function(x) x > 0.75)) %>%
  colSums()
```

And show \> 0.9

```{r}
ICC_treatment_2DG_normalization_merged %>%
  column_to_rownames("lipid") %>%
  mutate(across(everything(), function(x) x > 0.9)) %>%
  colSums()

ICC_cell_lines_normalization_merged %>%
  column_to_rownames("lipid") %>%
  mutate(across(everything(), function(x) x > 0.9)) %>%
  colSums()
```

### Write to file

Create with 1.5 times the size and letters and scale later.

```{r}
save_as_png_svg_pdf(
  "Figure3",
  all_plots_3_largeviolin,
  base_height = 1.5*page_width, base_width = 1.5*page_width)
```

## Figure 4

Figure 4 Correlation between class averages before normalization shows an overall agreement for structural lipids, but not for triglycerides and some signaling lipids.

Lipid abundances are log2 transformed, scaled, and centered. Class averages are calculated for cell line biological replicates before normalization, and Pearson correlation is used to correlate class averages over all samples. Rows and columns are clustered using hierarchical clustering with Euclidean distances and complete linkage.

Following instructions on <https://slowkow.com/notes/pheatmap-tutorial/> and <https://biocorecrg.github.io/CRG_RIntroduction/pheatmap-function-from-the-pheatmap-package.html>

```{r}
class_average_correlations <- read_csv(here(data_dir, "class_average_correlations.csv")) %>%
  rename("unknown" = "unknown.sphingolipid") %>%
  mutate(lipid_subclass = str_replace(lipid_subclass, "unknown.sphingolipid", "unknown")) %>%
  column_to_rownames("lipid_subclass")

min(class_average_correlations)
max(class_average_correlations)

# Sort dendogram to emphasize similarity
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
cluster_cols <- sort_hclust(hclust(dist(class_average_correlations)))
plot(cluster_cols, xlab = "", sub = "")

colnames(class_average_correlations) <- rownames(class_average_correlations)
fig4a <- pheatmap(class_average_correlations, 
         cluster_rows = cluster_cols,
         cluster_cols = cluster_cols,
         treeheight_row = 4*fontsize_smaller,
         treeheight_col = 4*fontsize_smaller,
         border_color = NA,
         fontsize = fontsize_larger,
         fontsize_row = fontsize_smaller,
         fontsize_col = fontsize_smaller,
         angle_col = "315")
fig4a
```

Include correlation of raw data not transformed sums of all structural lipids.

```{r}
plot_correlation_sums <- function(df, x, y){
  df <- df %>%
    rename_with(~str_replace_all(.x, "_", " "))
  cor_nr <- round(cor(df[[x]], df[[y]], method = "spearman"),2)
  print(cor_nr)
  
  p <- df %>%
    ggplot(aes(x = .data[[x]], y = .data[[y]])) +
    geom_point() +
    personal_theme() +
    geom_label(x = Inf, y = Inf, hjust = 1, vjust = 1,
              label = str_glue("r[s] == {cor_nr}"),
              label.size = NA, parse = T)
  return(p)
}
```

```{r fig.height=5, fig.width=3}
sums_of_SL_and_TGs <- read_csv(here(data_dir, "rawdata_sums_of_structural_lipids_and_TGs.csv")) %>%
  left_join(metadata, by = "Original_name") %>%
  rename("cell count (million)" = cell_nr_10E6)
head(sums_of_SL_and_TGs)

fig4b <- sums_of_SL_and_TGs %>%
  plot_correlation_sums(x = "cell count (million)", y = "structural lipid sum")

fig4c <- sums_of_SL_and_TGs %>%
  plot_correlation_sums(x = "cell count (million)", y = "TG sum")

plot_grid(fig4b, fig4c,ncol = 1)
```

#### HCC38

Add replicate plots to figure 4 to show replicate (dis)similarity

```{r}
plot_replicates_TG <- function(df_cell_line, repx, repy, show_legend = "none"){
  p <- ggplot(df_cell_line, aes(x = .data[[repx]], y = .data[[repy]], color = is_TG)) + 
    geom_point() +
    personal_theme() +
    scale_color_manual(values = list("TG" = "grey80", "not TG" = "grey20")) +
    geom_abline(slope = 1, intercept = 0, color = "blue") +
    theme(legend.position = show_legend) + labs(color = NULL)
  return(p)
}
# todo to calculations script
raw_data_HCC38 <- data_cell_lines_notScaled[["Not"]] %>%
    filter(str_detect(Original_name, "HCC38")) %>%
  left_join(select(metadata, cell_line, biological_replicate, Original_name),
            by = "Original_name") %>%
  unite(Original_name, cell_line, biological_replicate, sep = " rep ") %>%
  relocate(Original_name) %>%
  pivot_longer(!Original_name, names_to = "internal_identifiers", values_to = "log2 lipid") %>%
    pivot_wider(id_cols = internal_identifiers, names_from = "Original_name", values_from = "log2 lipid") %>%
  mutate(is_TG = ifelse(str_detect(internal_identifiers, "ATG_TG"), "TG", "not TG"))

fig4d <- plot_grid(
  plot_replicates_TG(raw_data_HCC38, "HCC38 rep 1", "HCC38 rep 2") + 
    theme(legend.position = "inside",legend.position.inside = c(0,0.9)),
  plot_replicates_TG(raw_data_HCC38, "HCC38 rep 1", "HCC38 rep 3"),
  plot_replicates_TG(raw_data_HCC38, "HCC38 rep 2", "HCC38 rep 3"),
  nrow = 1
)
fig4d
```

```{r fig.height=5.5, fig.width=5.5}
all_plots_4 <- plot_grid(
  plot_grid(fig4a$gtable, 
          plot_grid(fig4b, fig4c,
                   ncol= 1, labels=LETTERS[2:3]),
          nrow = 1, 
          labels = c(LETTERS[1],""),
          rel_widths = c(0.7,0.3)
  ), 
  fig4d, nrow = 2, rel_heights = c(2,1), labels = c("", LETTERS[4])
)
all_plots_4
```

### Write to file

```{r}
save_as_png_svg_pdf("Figure4", 
                    all_plots_4 + theme(plot.background = element_rect(fill = "white", color = "white")),
                    base_height = 12)
```

## Figure 5

Figure 5 Differences in PE class average between morphologies are partially caused by normalization.

(A) Lipid class averages for TG and PE before normalization of samples show a positive correlation (r=0.65). (B) After PQN normalization using all lipids, a negative correlation is observed (r=-0.76). (C) After PQN normalization with only structural lipids, the previously observed correlation is rescued to an extent (r=0.31). (D) Lower TG class averages are observed for mesenchymal cell lines and (E) higher overall PE after normalization using all lipids. PQN using structural lipids keeps the lower overall TG abundance in (F) mesenchymal cell lines but (G) does not artificially increase PE abundance. Class averages are the mean per class of log2 transformed autoscaled lipid abundances. E=epithelial, M=mesenchymal. Correlations are Pearson correlations.

### Load data

```{r}
class_averages_by_structural_lipids <- read_csv(here(data_dir, "class_averages_by_structural_lipids.csv")) 
class_averages_merged <- read_csv(here(data_dir, "class_averages_merged.csv")) 
class_averages_rawdata <- read_csv(here(data_dir, "class_averages_TGPE_rawdata.csv")) 

# For reviewer comment on autoscaling
class_averages_by_structural_lipids <- read_csv(here(data_dir, "noAutoscale_class_averages_by_structural_lipids.csv")) 
class_averages_merged <- read_csv(here(data_dir, "noAutoscale_class_averages_merged.csv")) 
class_averages_rawdata <- read_csv(here(data_dir, "noAutoscale_class_averages_TGPE_rawdata.csv")) 

```

### Violin plots

```{r}
violin_plot_lipid_class_average <- function(df){
  p <- metadata %>%
    dplyr::select(Original_name, Morphology) %>%
    unique() %>%
    right_join(df, by = "Original_name") %>%
    filter(Morphology %in% c("Epithelial-like", "Mesenchymal-like")) %>%
  ggplot(aes(x = Morphology, y = class_mean, fill = Morphology)) +
    geom_violin() +
    geom_boxplot(position=position_dodge(1), width=0.2, fill = "white") +
    scale_fill_manual(values = color_legend_morphology) +
    personal_theme() +
    labs(x = NULL, y = "class average")
  return(p)
}
```

```{r fig.height=6, fig.width=3}
TG_rawdata <- class_averages_rawdata %>% filter(class == "TG") %>%
  violin_plot_lipid_class_average()
PE_rawdata <- class_averages_rawdata %>% filter(class == "PE") %>%
  violin_plot_lipid_class_average()

TG_merged <- class_averages_merged %>% filter(class == "TG") %>%
  violin_plot_lipid_class_average()
PE_merged <- class_averages_merged %>% filter(class == "PE") %>%
  violin_plot_lipid_class_average()

TG_by_structural_lipids <- class_averages_by_structural_lipids %>% filter(class == "TG") %>%
  violin_plot_lipid_class_average()
PE_by_structural_lipids <- class_averages_by_structural_lipids %>% filter(class == "PE") %>%
  violin_plot_lipid_class_average()

plot_grid(
  plot_grid(TG_rawdata + ggtitle("raw data"), PE_rawdata, align = "h"),
  plot_grid(TG_merged + ggtitle("merged"), PE_merged, align = "h"),
  plot_grid(TG_by_structural_lipids + ggtitle("by structural lipids"), PE_by_structural_lipids, align = "h"),
  nrow = 3
)
```

### Correlation plots

```{r}
# correlation between the class averages after normalization
extract_TG_PE <- function(df){
  df <- df %>%
    filter(class %in% c("TG", "PE")) %>%
    dplyr::select(Original_name, class, class_mean) %>%
    pivot_wider(id_cols = Original_name, names_from = class, values_from = class_mean)
  return(df)
}
calculate_TG_PE_cor <- function(df){
  cor_nr <- round(cor(df$PE, df$TG), 2)
  return(cor_nr)
}

correlation_plot_TGPE <- function(class_averages, show_cor = F){
  cor_nr <- calculate_TG_PE_cor(class_averages)
  print(cor_nr)
  p <- class_averages %>%
    ggplot(aes(x = PE, y = TG)) +
      geom_point() +
      personal_theme() +
      geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") 
  if(show_cor){
    p <- p +
      geom_label(x = Inf, y = Inf, hjust = 1, vjust = 1,
              label = str_glue("r[s] == {cor_nr}"),
              label.size = NA, parse = T)
  }
  return(p)
}
```

```{r fig.height=2, fig.width=6}
# correlation between the class averages after structural lipid-based normalization

class_averages <- extract_TG_PE(class_averages_by_structural_lipids)
PE_TG_correlation_by_structural_lipids <- correlation_plot_TGPE(class_averages, show_cor = T)

# correlation between the class averages after merged normalization

class_averages <- extract_TG_PE(class_averages_merged)
PE_TG_correlation_merged <- correlation_plot_TGPE(class_averages, show_cor = T)

# correlation between the class averages before normalization
class_averages <- extract_TG_PE(class_averages_rawdata)
PE_TG_correlation_rawdata <- correlation_plot_TGPE(class_averages, show_cor = T)

plot_grid(PE_TG_correlation_rawdata, PE_TG_correlation_merged, PE_TG_correlation_by_structural_lipids, nrow = 1)
```

### Normalized HCC38

#### Load data

```{r}
# notScaled dataset
file_name <- here(data_dir, "data_normalized", "cell_lines", str_glue("cell_lines_SL_PQNNormalized_log2Transformed_notScaled.csv"))

if(!file.exists(file_name)) stop(str_glue("file not found: {file_name}"))
data_cell_lines_notScaled[["PQN_SL"]] <- read_csv(file_name, show_col_types = F)
```

```{r fig.height=2, fig.width=6}
get_HCC38 <- function(normalization_name){
  df <- data_cell_lines_notScaled[[normalization_name]] %>%
    filter(str_detect(Original_name, "HCC38")) %>%
    left_join(select(metadata, cell_line, biological_replicate, Original_name),
              by = "Original_name") %>%
    unite(Original_name, cell_line, biological_replicate, sep = " rep ") %>%
    relocate(Original_name) %>%
    pivot_longer(!Original_name, names_to = "internal_identifiers", values_to = "log2 lipid") %>%
      pivot_wider(id_cols = internal_identifiers, names_from = "Original_name", values_from = "log2 lipid") %>%
    mutate(is_TG = ifelse(str_detect(internal_identifiers, "ATG_TG"), "TG", "not TG"))
  return(df)
}
```

```{r fig.height=8, fig.width=5.5}
all_plots_5 <- plot_grid(
  # A-C
  PE_TG_correlation_rawdata + ggtitle("Not"),
  PE_TG_correlation_merged + ggtitle("PQN"), 
  PE_TG_correlation_by_structural_lipids + ggtitle("PQN SL"),
  # D-F
  TG_rawdata + labs(y = "TG"), 
  TG_merged + labs(y = "TG"), 
  TG_by_structural_lipids + labs(y = "TG"),
  # G-I
  PE_rawdata + labs(y = "PE"), 
  PE_merged + labs(y = "PE"),
  PE_by_structural_lipids + labs(y = "PE"),
  # J-L
  plot_replicates_TG(get_HCC38("Not"), "HCC38 rep 1", "HCC38 rep 2") + 
    theme(legend.position = "inside",legend.position.inside = c(0,0.9)),
  plot_replicates_TG(get_HCC38("PQN"), "HCC38 rep 1", "HCC38 rep 2"),
  plot_replicates_TG(get_HCC38("PQN_SL"), "HCC38 rep 1", "HCC38 rep 2"),
  nrow = 4,
  labels = LETTERS[1:12]
)
all_plots_5
```

### Write to file

```{r}
save_as_png_svg_pdf(
  "Figure5",
  all_plots_5,
  base_height = 16)
```

## Figure 6

Figure 6 Correlation of PCs after data transformation shows similar results.

The (A) first, (B) second, and (C) third principal components after log2 or cube root transformation for lipids measured in cell lines show similar contributions of individual lipids. Grey represents the x=y. Each dot is the loading value of a lipid. Lipid classes are represented in different colors. Lipid abundances were scaled using autoscaling before performing PCA. (D) Overlap of differential abundant lipids between epithelial and mesenchymal morphologies without transformation or after log2 or cube root transformation.

### Load data

```{r}
pqn_data_log2 <- read_csv(here(data_dir, "pqn_data_log2.csv")) %>% column_to_rownames("group")
pqn_data_cuberoot <- read_csv(here(data_dir, "pqn_data_cuberoot.csv")) %>% column_to_rownames("group")
```

### Plot PCA loadings

```{r}
make_comparison_plot <- function(PCA_comparison, x, y, multiplyPC){
  p <- PCA_comparison %>%
    left_join(lipid_info, by = c("lipid" = "internal_identifiers")) %>%
    ggplot(aes(x = .data[[x]], y = .data[[y]])) +#, color = class)) +
    geom_point(color = "grey20") +
    personal_theme() +
    labs(x = str_replace_all(x, "_", " "), 
         y = str_replace_all(y, "_", " ")) +
    geom_abline(slope = 1*multiplyPC, color = "black", linetype = "dashed")
  return(p)
}

plot_loadings <- function(pca_log2, pca_cuberoot, show_positivePC1 = T, show_positivePC2 = T, 
                          plot_PC3 = F, show_positivePC3 = T){
  # Set by hand whether I want to flip a PC axis
  multiplyPC1 <- ifelse(show_positivePC1, 1, -1)
  multiplyPC2 <- ifelse(show_positivePC2, 1, -1)
  
  # Plot correlation between loading values
  PCA_comparison <- full_join(
    data.frame(pca_log2$loadings$X) %>% rownames_to_column("lipid"),
    data.frame(pca_cuberoot$loadings$X) %>% rownames_to_column("lipid"),
    by = "lipid",
    suffix = c("_log2", "_cuberoot")
  )
  
  # Compare correlation between first principal components for both modes
  p_pca1 <- make_comparison_plot(PCA_comparison, x = "PC1_log2", y = "PC1_cuberoot", multiplyPC1)
  
  # Compare correlation between second principal components for both modes
  p_pca2 <- make_comparison_plot(PCA_comparison, x = "PC2_log2", y = "PC2_cuberoot", multiplyPC2)
  
  if(plot_PC3){
    multiplyPC3 <- ifelse(show_positivePC3, 1, -1)
    # Compare correlation between second principal components for both modes
    p_pca3 <- make_comparison_plot(PCA_comparison, x = "PC3_log2", y = "PC3_cuberoot", multiplyPC3)
    return(list(p_pca1, p_pca2, p_pca3))
  }else{
    return(list(p_pca1, p_pca2))
  }
}
```

```{r}
pca_log2 <- mixOmics::pca(pqn_data_log2, scale = F, ncomp = 3)
pca_cuberoot <- mixOmics::pca(pqn_data_cuberoot, scale = F, ncomp = 3)
print(str_c("Overlapping nr of max lipids in PC1 is ", 
  intersect(
    pca_log2$loadings$X %>% data.frame() %>% arrange(desc(abs(PC1))) %>% rownames() %>% head(10),
    pca_cuberoot$loadings$X %>% data.frame() %>% arrange(desc(abs(PC1))) %>% rownames() %>% head(10)) %>%
  length()))
pca_plots <- plot_loadings(pca_log2, pca_cuberoot, show_positivePC1 = T, show_positivePC2 = T,
                           plot_PC3 = T, show_positivePC3 = T)

all_plots <- plot_grid(
  pca_plots[[1]] + theme(legend.position = "none"), 
  pca_plots[[2]] + theme(legend.position = "none"),
  pca_plots[[3]] + theme(legend.position = "none"),
  align = 'vh',
  labels = LETTERS[1:3],
  hjust = -1,
  nrow = 1
)

all_plots
```

### Venn diagram after differential analysis

#### Load data

```{r}
load(here(data_dir, "venn_diagram_differential_lipids.RData"))
```

```{r fig.height=5/2.54, fig.width=14/4/2.54}
venn_diagram <- ggvenn(venn_diagram_differential_lipids, 
       show_percentage = F, fill_alpha = 0,
       stroke_size = dot_scaling_size,
       set_name_size = fontsize_larger/3, text_size = fontsize_larger/4) +
  theme(plot.background = element_rect(fill = NA, color = NA)) +
  coord_cartesian(clip="off")
venn_diagram
```

### Combine plots

```{r fig.height=2, fig.width=5.5}
plot_grid(NULL, venn_diagram, NULL, rel_heights = c(0.1,1,0.1))

all_plots_6 <- plot_grid(
  pca_plots[[1]], 
  pca_plots[[2]], 
  pca_plots[[3]], 
  plot_grid(NULL, venn_diagram, NULL, rel_heights = c(0.2,1,0.2), ncol = 1), 
  labels = LETTERS[1:4],
  #hjust = -1,
  nrow = 1
)

all_plots_6
```

### Write to file

```{r fig.height=2, fig.width=5.5}
save_as_png_svg_pdf(
  "Figure6",
  all_plots_6 + theme(plot.background = element_rect(fill = "white", color = "white")),
  base_height = 4.5)
```

# Supplementary figures

## Supplementary Figure 1

```{r fig.height=4, fig.width=6}
all_plots_sup1 <- plot_grid(
  plot_grid(plotlist = list(
  plot_two_ICC("Sum", "Median"),
  plot_two_ICC("Sum", "PQN"),
  plot_two_ICC("PQN", "Median")),
  nrow = 1, labels = LETTERS[1:3]))
all_plots_sup1
```

### Write to file

```{r fig.height=8, fig.width=5.5}
save_as_png_svg_pdf(
  "SupplementaryFigure1",
  all_plots_sup1,
  base_height = 5)
```

## Supplementary Figure 2

2DG supplementary

```{r fig.height=4, fig.width=6}
plot_list_ICC_treatment_2DG <- lapply(normalization_order, function(name) plot_two_ICC("Not", name, df = ICC_treatment_2DG_normalization_merged) + ggtitle(name))
icc_overlaps_treatment_2DG <- plot_ICC(all_ICCs = ICC_treatment_2DG_normalization_merged)
plot_list_ICC_treatment_2DG[1] <- icc_overlaps_treatment_2DG[2]

all_plots_sup2 <- cowplot::plot_grid(plotlist = plot_list_ICC_treatment_2DG, 
                     nrow = 1, labels = LETTERS[1:2], label_size = 2*figure_label_size,
                     vjust = 1, hjust = 0)
all_plots_sup2
```

### Write to file

```{r fig.height=8, fig.width=5.5}
save_as_png_svg_pdf(
  "SupplementaryFigure2",
  all_plots_sup2,
  base_height = 6, base_width = 2*page_width)
```

## Supplementary Figure 3

```{r fig.height=5, fig.width=3}
figsup3a <- sums_of_SL_and_TGs %>%
  rename("prot conc. (mg/mL)" = corrected_conc_mg_per_mL) %>%
  plot_correlation_sums(x = "prot conc. (mg/mL)", y = "structural lipid sum")

figsup3b <- sums_of_SL_and_TGs %>%
  rename("prot conc. (mg/mL)" = corrected_conc_mg_per_mL) %>%
  plot_correlation_sums(x = "prot conc. (mg/mL)", y = "TG sum")

plot_grid(figsup3a, figsup3b ,ncol = 1)
```

```{r fig.height=2.6, fig.width=5.5}
all_plots_sup3 <- plot_grid(
  figsup3a,
  figsup3b,
  align = 'vh',
  labels = LETTERS[1:2],
  hjust = -1,
  nrow = 1
)
all_plots_sup3
```

### Write to file

```{r fig.height=8, fig.width=5.5}
save_as_png_svg_pdf(
  "SupplementaryFigure3",
  all_plots_sup3,
  base_height = 5)
```

## Supplementary Figure 4

ICCs after log2 or cube root transformation

```{r}
ICC_cell_lines_transformation <- read_csv(here(data_dir, "ICC_cell_lines_transformation.csv"))
```

```{r fig.height=2, fig.width=5.5}

all_plots_sup4 <- plot_grid(
  plot_two_ICC("not", "log2", df = ICC_cell_lines_transformation),
  plot_two_ICC("not", "cuberoot", df = ICC_cell_lines_transformation),
  plot_two_ICC("log2", "cuberoot", df = ICC_cell_lines_transformation),
  labels = LETTERS[1:3],
  #hjust = -1,
  nrow = 1
)
all_plots_sup4
```

### Write to file

```{r fig.height=2.6, fig.width=5.5}
save_as_png_svg_pdf(
  "SupplementaryFigure4",
  all_plots_sup4,
  base_height = 5
)
```

## Supplementary Figure 5

```{r fig.height=3.15, fig.width=6.4}
BT549_density_info <- read_csv(here(data_dir, "BT549_density_info.csv")) %>%
  rename(replicate = biological_replicate)
```

```{r fig.height=3.15, fig.width=6.4}
plot_BT549_density_info <- function(df,y_label = F){
  df <- mutate(df, normalization = as.factor(str_replace(normalization, "Normalized", " normalized"))) %>%
    mutate(replicate = factor(replicate))
  p1 <- df %>%
    filter(transformation == "log2Transformed") %>%
  ggplot(aes(x = value, color = replicate)) + geom_density() +
    facet_grid(~ normalization, scales = "free") +
    personal_theme() +
    theme(strip.background =element_rect(fill="white")) +
    labs(x = NULL, y = NULL)
  
  p2 <- df %>%
    filter(transformation != "log2Transformed") %>%
  ggplot(aes(x = value, color = replicate)) + geom_density() +
    facet_grid(~ normalization, scales = "free") +
    personal_theme() +
    theme(strip.text = element_blank()) +
    labs(x = NULL, y = NULL) +
    theme(legend.position = "inside",legend.position.inside = c(0.92,0.9))
  if(y_label){
    p1 <- p1 + labs(x = NULL, y = "log2 transformed")
    p2 <- p2 + labs(x = NULL, y = "cube root transformed")
  }
  return(plot_grid(p1, p2, nrow = 2, align = "vh"))
}
p_sup5 <- BT549_density_info %>% 
  plot_BT549_density_info(y_label = T)
p_sup5
```

```{r fig.height=3.15, fig.width=6.4}
all_plots_sup5 <- plot_grid(
  p_sup5,
  align = 'vh',
  hjust = -1,
  nrow = 1
)

all_plots_sup5
```

### Write to file

```{r fig.height=3.15, fig.width=6.4}
save_as_png_svg_pdf(
  "SupplementaryFigure5",
  all_plots_sup5 + theme(plot.background = element_rect(fill = "white", color = "white")),
  base_height = 8)
```

# Session info

```{r}
sessionInfo()
```
