---
title: "Calculations to get all data for all figures"
author: "Hanneke Leegwater"
date: "`r Sys.Date()`"
engine: knitr
format: 
  html:
    self-contained: true
    toc: true
    toc_depth: 4
    toc_float: true
    df-print: kable
---

## Introduction

This Quarto script reproduces all calculations performed for the manuscript "Distinct lipidomic profiles in breast cancer cell lines relate to proliferation and EMT phenotypes".

## Preparations

### Load R libraries

```{r}
library(tidyverse)
library(writexl)
library(here)
library(limma)
library(DESeq2)
library(gprofiler2)
library(AnnotationDbi)
library(org.Hs.eg.db)

# Custom functions
source(here("code/functions/lipid_overrepresentation.R"))
source(here("code/functions/calculate_lipid_class_average.R"))
source(here("code/functions/duplicate_control_T0_samples.R"))
source(here("code/functions/scale_lipid_to_T0is0.R"))
```

### Define location of data

The raw data can be found in data/biology_manuscript_files.

```{r}
data_dir <- here("data/biology_manuscript_files")
str_glue("Using data from {data_dir}")
stopifnot(dir.exists(data_dir))
```

### Create output directory

```{r}
datestamp <- str_remove_all(Sys.Date(), "-")
output_dir <- here("manuscript_biology", str_glue("{datestamp}_data_for_figures"))
str_glue("Output directory path is {output_dir}")
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
output_dir
```

### Define how to save files

Change write_csv so that it always writes na as "".
And print output file path.

```{r}
write_csv <- function(data, file, ...) {
  readr::write_csv(data, file, na = "", ...)
  print(stringr::str_glue("Dataset saved to {file}"))
}
```

## Open metadata

```{r}
# Sample metadata
metadata <- read_csv(here(data_dir, "metadata_cell_lines.csv")) %>%
  mutate(Morphology_short = str_extract(Morphology, "^\\w", )) %>%
  mutate(Morphology = str_replace(Morphology, "_", "-"))

print(dim(metadata))
head(metadata)
write_csv(metadata, here(output_dir, "metadata.csv"))

# Information on lipid classes
lipid_info <- read_csv(here(data_dir, "lipid_info.csv"))

print(dim(lipid_info))
head(lipid_info)
write_csv(lipid_info, here(output_dir, "lipid_info.csv"))
```

## Figures

### Figure 1

#### Load data

```{r}
data_cell_lines <- read_csv(here(data_dir, "raw_counts_2DRNAseq_celllines_20250212.csv"))
gene_identifier_table <- read_csv(here(data_dir, "gene_identifier_table_unique.csv"),
  col_types = cols(NCBI_Gene_ID = col_character())
)
```

#### Get GO metabolic process

```{r}
lipid_metabolic_process <- AnnotationDbi::select(org.Hs.eg.db,
  keys = "GO:0006629", columns = c("SYMBOL"), keytype = "GOALL"
) %>%
  dplyr::select(GOALL, SYMBOL) %>%
  distinct() %>%
  dplyr::rename(gene_symbol = SYMBOL)
write_csv(lipid_metabolic_process, here(output_dir, "lipid_metabolic_process_GO.0006629.csv"))
```

#### Run differential expression analysis

Prepare data: keep only genes that have more than 1 cpm in at least 5 cell lines.
Amount of million reads per sample is:

```{r}
data_cell_lines %>%
  dplyr::select(-EnsemblID) %>%
  colSums() %>%
  lapply(function(x) round(x / 1E6, 1)) %>%
  unlist() %>%
  sort()
```

So approximately 1 CPM is about at least 10 counts.
Keep genes that:

-   have a count \> 10 in 5 or more cell lines, and
-   have a mean count more than 1.

```{r}
## Prepare columns and rows
column_info <- metadata %>%
  dplyr::select(group, Morphology) %>%
  unique() %>%
  filter(Morphology != "treatment_study") %>%
  filter(group %in% colnames(data_cell_lines)) %>%
  mutate(Morphology = str_replace_all(Morphology, "-", "_")) %>%
  mutate(Morphology = factor(Morphology))
dim(column_info)
head(column_info)

all_counts <- column_to_rownames(data_cell_lines, "EnsemblID") %>%
  dplyr::select(all_of(column_info$group)) %>%
  filter(rowSums(. > 10) > 5)
dim(all_counts)
all_counts <- all_counts %>%
  filter(rowMeans(.) > 1)
dim(data_cell_lines)
dim(all_counts)
head(all_counts)

## order of the metadata file has to be the same as the count file. Check if this is true:
identical(column_info$group, colnames(all_counts))
stopifnot(identical(column_info$group, colnames(all_counts)))
```

Perform differential expression analysis:

```{r}
dds <- DESeqDataSetFromMatrix(
  countData = all_counts,
  colData = column_info,
  design = ~Morphology
)

dds <- estimateSizeFactors(dds)
dds <- DESeq(dds)
```

```{r}
file_name <- here(output_dir, "dds_object_Morphology.RData")
print(str_glue("Saving deseq object to {file_name}"))
save(dds, file = file_name)

## Extract normalized counts
all_counts_norm <- as.data.frame(counts(dds, normalized = TRUE)) %>%
  rownames_to_column("EnsemblID")

dplyr::select(gene_identifier_table, NCBI_Gene_ID, Ensembl_gene_ID, gene_symbol) %>%
  right_join(all_counts_norm, by = c("Ensembl_gene_ID" = "EnsemblID")) %>%
  write_csv(here(output_dir, "normalized_counts.csv"))

all_counts_norm_log2 <- all_counts_norm %>%
  left_join(
    dplyr::select(gene_identifier_table, Ensembl_gene_ID, gene_symbol),
    by = c("EnsemblID" = "Ensembl_gene_ID")
  ) %>%
  dplyr::select(-EnsemblID) %>%
  mutate(across(where(is.numeric), function(x) log2(x + 1)))

all_counts_norm_log2 %>%
  write_csv(here(output_dir, "all_counts_norm_log2.csv"))
```

Study the results and print statistics:

```{r}
resultsNames(dds) # lists the coefficients

lipid_DEGs <- results(dds, name = "Morphology_Mesenchymal_like_vs_Epithelial_like") %>%
  data.frame() %>%
  arrange(padj) %>%
  rownames_to_column("EnsemblID") %>%
  left_join(
    dplyr::select(
      gene_identifier_table, Ensembl_gene_ID, gene_symbol,
      NCBI_Gene_ID, HGNC_ID, gene_name
    ),
    by = c("EnsemblID" = "Ensembl_gene_ID")
  ) %>%
  filter(gene_symbol %in% lipid_metabolic_process$gene_symbol)

write_csv(lipid_DEGs, here(output_dir, "lipid_DEGs_MvsE.csv"))
```

Information on metadata for in the manuscript:

```{r}
dim(column_info)

table(column_info$Morphology)
```

```{r}
str_glue("All genes in list of GO:0006629 lipid metabolic process: {nrow(lipid_metabolic_process)}")
str_glue("Total nr of genes studied in differential gene expression analysis: {nrow(lipid_DEGs)}")
str_glue("Higher in mesenchymal-like, logFC > 0: {nrow(filter(lipid_DEGs, log2FoldChange > 0))}")
str_glue("Higher in mesenchymal-like, logFC > 1: {nrow(filter(lipid_DEGs, log2FoldChange > 1))}")
str_glue("Higher in mesenchymal-like, logFC > 0 & padj < 0.05: {nrow(filter(lipid_DEGs, log2FoldChange >0, padj < 0.05))}")
str_glue("Higher in mesenchymal-like, logFC > 0 & padj < 0.01: {nrow(filter(lipid_DEGs, log2FoldChange >0, padj < 0.01))}")
str_glue("Higher in mesenchymal-like, logFC > 1 & padj < 0.05: {nrow(filter(lipid_DEGs, log2FoldChange >1, padj < 0.05))}")
str_glue("Higher in mesenchymal-like, logFC > 1 & padj < 0.01: {nrow(filter(lipid_DEGs, log2FoldChange >1, padj < 0.01))}")

str_glue("Higher in epithelial-like, logFC > 0: {nrow(filter(lipid_DEGs, log2FoldChange < 0))}")
str_glue("Higher in epithelial-like, logFC > 1: {nrow(filter(lipid_DEGs, log2FoldChange < -1))}")
str_glue("Higher in epithelial-like, logFC > 0 & padj < 0.05: {nrow(filter(lipid_DEGs, log2FoldChange < 0, padj < 0.05))}")
str_glue("Higher in epithelial-like, logFC > 0 & padj < 0.01: {nrow(filter(lipid_DEGs, log2FoldChange < 0, padj < 0.01))}")
str_glue("Higher in epithelial-like, logFC > 1 & padj < 0.05: {nrow(filter(lipid_DEGs, log2FoldChange < -1, padj < 0.05))}")
str_glue("Higher in epithelial-like, logFC > 1 & padj < 0.01: {nrow(filter(lipid_DEGs, log2FoldChange < -1, padj < 0.01))}")
```

Check CD36.
In this dataset, values below 10 are "lower than 1 CPM".

```{r}
lipid_DEGs %>%
  filter(gene_symbol == "CD36")

all_counts_norm %>%
  filter(EnsemblID == "ENSG00000135218") %>%
  pivot_longer(!EnsemblID) %>%
  mutate(more_than_1_CPM = ifelse(value > 10, "yes", "no")) %>%
  pull(more_than_1_CPM) %>%
  table()
```

#### Perform enrichment analysis

See https://cran.r-project.org/web/packages/gprofiler2/vignettes/gprofiler2.html#gene-list-functional-enrichment-analysis-with-gost for details

```{r}
lipid_DEGs_more_M <- lipid_DEGs %>%
  filter(padj < 0.01, log2FoldChange > 1) %>%
  filter(!is.na(gene_symbol)) %>%
  pull(gene_symbol)
length(lipid_DEGs_more_M)

lipid_DEGs_more_E <- lipid_DEGs %>%
  filter(padj < 0.01, log2FoldChange < -1) %>%
  filter(!is.na(gene_symbol)) %>%
  pull(gene_symbol)
length(lipid_DEGs_more_E)
lipid_DEGs_more_E
```

The next code chunk requires an internet connection.

```{r}
enrichment_results <- gost(
  list(
    "more_M" = lipid_DEGs_more_M,
    "more_E" = lipid_DEGs_more_E
  ),
  organism = "hsapiens", ordered_query = TRUE, correction_method = "gSCS",
  evcodes = TRUE
)

save(enrichment_results,
  file = here(output_dir, str_glue("gprofiler_enrichment_object.RData"))
)
write_csv(
  enrichment_results$result,
  here(output_dir, "gprofiler_enrichment_results.csv")
)
```

```{r}
head(enrichment_results$result)
```

```{r}
names(enrichment_results$meta)

names(enrichment_results$meta)
enrichment_results$meta$version
enrichment_results$meta$timestamp

names(enrichment_results$meta$result_metadata)
names(enrichment_results$meta$query_metadata)
```

```{r}
# Compare E and M
intersect(
  filter(enrichment_results$result, query == "more_E") %>% pull(term_name),
  filter(enrichment_results$result, query == "more_M") %>% pull(term_name)
) %>%
  head(10)
# setdiffs
setdiff(
  filter(enrichment_results$result, query == "more_E") %>% pull(term_name),
  filter(enrichment_results$result, query == "more_M") %>% pull(term_name)
) %>%
  head(10)
setdiff(
  filter(enrichment_results$result, query == "more_M") %>% pull(term_name),
  filter(enrichment_results$result, query == "more_E") %>% pull(term_name)
) %>%
  head(10)

enrichment_results$result %>%
  filter(query == "more_M") %>%
  filter(str_detect(term_name, "phospholipid")) %>%
  dplyr::select(
    term_name, term_id, term_size, query_size, intersection_size,
    p_value, precision
  )

enrichment_results$result %>%
  arrange(p_value) %>%
  filter(term_size < 10) %>%
  head(10)
```

Print info on specific pathways

```{r}
GO_names <- c(
  "phospholipid metabolic process",
  "fatty acid metabolic process",
  "lipid transport",
  "sphingolipid metabolic process"
)
enrichment_results$result %>%
  filter(term_name %in% GO_names) %>%
  relocate(
    term_name, query, precision,
    term_size, intersection_size
  ) %>%
  dplyr::select(-evidence_codes) %>%
  arrange(term_name)
```

```{r}
GO_lookup <- enrichment_results$result %>%
  dplyr::select(term_name, term_id) %>%
  filter(term_name %in% GO_names) %>%
  distinct()
GO_lookup
```

```{r}
results_GO <- AnnotationDbi::select(org.Hs.eg.db,
  keys = GO_lookup$term_id,
  columns = c("SYMBOL"), keytype = "GOALL"
) %>%
  dplyr::select(GOALL, SYMBOL) %>%
  distinct()
write_csv(results_GO, here(output_dir, "GO_terms_genes.csv"))
```

```{r}
pheatmap_annotation <- left_join(dplyr::select(lipid_DEGs, gene_symbol),
  (results_GO %>%
    mutate(included = 1) %>%
    pivot_wider(id_cols = SYMBOL, names_from = GOALL, values_from = included)
  ),
  by = c("gene_symbol" = "SYMBOL")
) %>%
  column_to_rownames("gene_symbol") %>%
  mutate(across(starts_with("GO"), ~ replace_na(., 0))) %>%
  mutate(across(starts_with("GO"), ~ factor(., levels = c("0", "1"))))

head(pheatmap_annotation)

pheatmap_annotation_n <- pheatmap_annotation %>%
  dplyr::rename(all_of(setNames(GO_lookup$term_id, GO_lookup$term_name)))
head(pheatmap_annotation_n, 20)
```

Save the annotation

```{r}
lipid_genes_log2_FC1_padj_0.01 <- lipid_DEGs %>%
  filter(abs(log2FoldChange) > 1, padj < 0.01) %>%
  arrange(padj) %>%
  relocate(gene_symbol) %>%
  left_join(
    pheatmap_annotation_n %>%
      rownames_to_column("gene_symbol"),
    by = "gene_symbol"
  )
lipid_genes_log2_FC1_padj_0.01 %>%
  dplyr::select(gene_symbol, all_of(colnames(pheatmap_annotation_n))) %>%
  write_csv(here(output_dir, "GO_terms_annotation.csv"))
```

#### Supplementary Table 3

Save the differentially expressed lipid-related gene list as supplementary file.
Also save the matching normalized counts values.

```{r}
lipid_genes_log2_FC1_padj_0.01 %>%
  dplyr::select(gene_symbol) %>%
  left_join(all_counts_norm_log2, by = "gene_symbol") %>%
  write_csv(here(output_dir, "lipid_genes_log2_FC1_padj_0.01.csv"))

write_csv(
  lipid_genes_log2_FC1_padj_0.01,
  here(output_dir, "Supplementary_Table_3_lipid_DEGs.csv")
)
here(output_dir, "Supplementary_Table_3_lipid_DEGs.xlsx")
writexl::write_xlsx(
  list(lipid_DEGs = lipid_genes_log2_FC1_padj_0.01),
  path = here(output_dir, "Supplementary_Table_3_lipid_DEGs.xlsx"),
  col_names = TRUE,
  format_headers = TRUE
)
```

### Figure 2

#### Load data

```{r}
data_cell_lines_autoScaled <- read_csv(here(data_dir, "cell_lines_replicateAveraged_PQNNormalized_log2Transformed_autoScaled.csv"))
data_cell_lines_autoScaled[1:6, 1:6]
```

```{r}
calculate_lipid_class_average <- function(df, lipid_info, subclass = F) {
  rename_first_col <- F
  if (colnames(df)[1] != "Original_name") {
    warning("The first column name is not Original_name. Are you sure this contains sample names?")
    first_col_name <- colnames(df)[1]
    rename_first_col <- T
    colnames(df)[1] <- "Original_name"
  }

  if (subclass == F) {
    df_average <- df %>%
      pivot_longer(!Original_name, names_to = "internal_identifiers", values_to = "value") %>%
      left_join(dplyr::select(lipid_info, internal_identifiers, class), by = "internal_identifiers") %>%
      group_by(Original_name, class) %>%
      summarise(
        class_mean = mean(value),
        class_sd = sd(value)
      ) %>%
      ungroup()
  } else {
    df_average <- df %>%
      pivot_longer(!Original_name, names_to = "internal_identifiers", values_to = "value") %>%
      left_join(dplyr::select(lipid_info, internal_identifiers, subclass), by = "internal_identifiers") %>%
      group_by(Original_name, subclass) %>%
      summarise(
        subclass_mean = mean(value),
        subclass_sd = sd(value)
      ) %>%
      ungroup()
  }

  if (rename_first_col) colnames(df_average)[1] <- first_col_name

  return(df_average)
}

lipid_class_averages_long <- data_cell_lines_autoScaled %>%
  calculate_lipid_class_average(lipid_info, subclass = T)

dim(lipid_class_averages_long)
head(lipid_class_averages_long)

lipid_class_averages <- lipid_class_averages_long %>%
  dplyr::select(-subclass_sd) %>%
  pivot_wider(id_cols = group, names_from = subclass, values_from = subclass_mean) %>%
  column_to_rownames("group")


class_average_correlations <- lipid_class_averages %>%
  cor() %>%
  data.frame()

colnames(class_average_correlations) <- rownames(class_average_correlations)

head(class_average_correlations)

class_average_correlations %>%
  rownames_to_column("lipid") %>%
  write_csv(here(output_dir, "lipid_class_average_correlations_after_SL_normalization.csv"))
```

#### Principal component analysis

```{r}
pca_lipid_classes <- lipid_class_averages %>%
  mixOmics::pca(scale = F, ncomp = 4)

PCs <- data.frame(pca_lipid_classes$variates$X) %>%
  rownames_to_column("group") %>%
  left_join(
    metadata %>%
      dplyr::select(group, Morphology, classification, Subtype, MigrationSpeed, ProliferationRate),
    by = "group"
  )

loadings <- pca_lipid_classes$loadings$X %>%
  data.frame() %>%
  rownames_to_column("lipid_class")

write_csv(PCs, here(output_dir, "PCA_variates_lipid_classes.csv"))
save(pca_lipid_classes, file = here(output_dir, "PCA_lipid_classes.RData"))
write_csv(loadings, here(output_dir, "PCA_loadings_lipid_classes.csv"))
```

Correlate metadata and principal components using spearman correlations

Using a tip from https://www.biostars.org/p/9527172/ for binarizing categorical variables.

```{r}
metadata_names <- data.frame(
  metadata = c(
    "Epithelial-like", "Mesenchymal-like", "Round", "Suspension", "Unknown",
    "NonTNBC", "TNBC",
    "BasalA", "BasalB", "Luminal",
    "MigrationSpeed", "ProliferationRate", "EMT_score"
  ),
  metadata_name = c(
    "Epithelial", "Mesenchymal", "Round", "Suspension", "Unknown",
    "HR+", "TNBC",
    "Basal A", "Basal B", "Luminal",
    "Migration", "Proliferation", "EMT score"
  ),
  category = c(
    rep("Morphology", 5),
    rep("Classification", 2),
    rep("Subtype", 3),
    rep("", 3)
  )
)

vector_to_binary_df <- function(vec) {
  df <- data.frame(samples = 1:length(vec), vals = vec) %>%
    pivot_longer(!samples) %>%
    pivot_wider(names_from = c(name, value), values_fn = length, values_fill = 0) %>%
    dplyr::select(-samples) %>%
    rename_with(~ str_remove(.x, "vals_"))
  return(df)
}

# Binarize all categorical variables
heatmap_metadata_data <- bind_cols(
  PCs,
  vector_to_binary_df(PCs$Morphology),
  vector_to_binary_df(PCs$classification),
  vector_to_binary_df(PCs$Subtype)
) %>%
  dplyr::select(-Morphology, -classification, -Subtype) %>%
  column_to_rownames("group") %>%
  cor(method = "spearman", use = "pairwise.complete.obs") %>%
  data.frame() %>%
  dplyr::select(PC1, PC2, PC3, PC4) %>%
  rownames_to_column("metadata") %>%
  filter(!metadata %in% c("PC1", "PC2", "PC3", "PC4")) %>%
  left_join(metadata_names, by = "metadata") %>%
  relocate(metadata_name) %>%
  dplyr::select(-metadata) %>%
  arrange(category)

heatmap_metadata_data %>%
  arrange(PC1) %>%
  head()

write_csv(heatmap_metadata_data, here(output_dir, "heatmap_metadata_data.csv"))
```

```{r}
lipid_class_averages_long %>%
  write_csv(here(output_dir, "lipid_class_averages_long_after_SL_normalization.csv"))
lipid_class_averages %>%
  rownames_to_column("cell_line") %>%
  write_csv(here(output_dir, "lipid_class_averages_after_SL_normalization.csv"))
```

#### Correlate gene expression with principal components

Only for cell lines that have matching transcriptome data.

```{r}
column_info <- PCs %>%
  dplyr::select(group, PC1, PC2, PC3, PC4) %>%
  filter(group %in% colnames(all_counts))

## order of the metadata file has to be the same as the count file. Check if this is true:
all_counts <- dplyr::select(all_counts, all_of(column_info$group))
stopifnot(identical(column_info$group, colnames(all_counts)))

run_deseq2 <- function(PC_name) {
  # Run DESeq2 with principal component value as design.
  dds <- DESeqDataSetFromMatrix(
    countData = all_counts,
    colData = column_info,
    design = as.formula(str_glue("~{PC_name}"))
  )
  dds <- DESeq(dds)

  # Add spearman correlations between gene expression and principal component
  gene_expression_correlations <- data.frame(
    col = apply(all_counts, 1, function(x) cor(x, column_info[[PC_name]], method = "spearman"))
  ) %>%
    dplyr::rename(!!str_glue("rs_{PC_name}") := col) %>%
    rownames_to_column("EnsemblID")

  # Format results and add gene name information
  results_df <- results(dds) %>%
    data.frame() %>%
    arrange(padj) %>%
    rownames_to_column("EnsemblID") %>%
    left_join(
      dplyr::select(
        gene_identifier_table, Ensembl_gene_ID, gene_symbol,
        NCBI_Gene_ID, HGNC_ID, gene_name
      ),
      by = c("EnsemblID" = "Ensembl_gene_ID")
    ) %>%
    left_join(gene_expression_correlations, by = "EnsemblID")
  return(results_df)
}

genes_matching_PC1 <- run_deseq2("PC1")
dim(genes_matching_PC1)
head(genes_matching_PC1)

genes_matching_PC2 <- run_deseq2("PC2")
dim(genes_matching_PC2)
head(genes_matching_PC2)

write_csv(genes_matching_PC1, here(output_dir, "genes_matching_PC1.csv"))
write_csv(genes_matching_PC2, here(output_dir, "genes_matching_PC2.csv"))
```


### Figure 3

Calculate an ANOVA per lipid class between basal A, basal B, and luminal cell lines.

```{r}
nr_classes <- length(unique(lipid_class_averages_long$subclass))
anova_info_lipid_class <- data.frame(
  subclass = unique(lipid_class_averages_long$subclass),
  pval_anova = rep(1, nr_classes),
  adj.r.squared = rep(0, nr_classes)
) %>% column_to_rownames("subclass")
for (sub_class in unique(lipid_class_averages_long$subclass)) {
  print(sub_class)
  df_subclass <- filter(lipid_class_averages_long, subclass == sub_class) %>%
    left_join(
      dplyr::select(metadata, group, Subtype) %>% unique(),
      by = "group"
    ) %>%
    dplyr::select(group, subclass_mean, Subtype)
  result <- lm(subclass_mean ~ Subtype, data = df_subclass)
  result_summary <- summary(result)
  print(result_summary)

  # Store adjusted Rsquared of the fit
  anova_info_lipid_class[sub_class, "adj.r.squared"] <- result_summary$adj.r.squared

  # Store pvalue related to subtypes
  result_anova <- anova(result)
  anova_info_lipid_class[sub_class, "pval_anova"] <- result_anova$`Pr(>F)`[[1]]
}

## Add multiple testing correction
anova_info_lipid_class$padj <- p.adjust(anova_info_lipid_class$pval_anova, method = "BH")
anova_info_lipid_class$padj_bonferroni <- p.adjust(anova_info_lipid_class$pval_anova, method = "bonferroni")
anova_info_lipid_class %>%
  arrange(padj) %>%
  mutate(across(adj.r.squared, ~ round(., 2))) %>%
  mutate(across(matches("pval|padj"), ~ round(., 4))) %>%
  head(n = 22)

anova_info_lipid_class %>%
  rownames_to_column("subclass") %>%
  arrange(padj) %>%
  write_csv(here(output_dir, "anova_results_lipid_class_and_Subtype.csv"))
```

### Figure 4

#### Load data

```{r}
data_cell_lines_notScaled <- read_csv(here(data_dir, "cell_lines_replicateAveraged_PQNNormalized_log2Transformed_notScaled.csv"))
```

#### Figure 4A-C

```{r}
## Filter to keep only cell lines in the lipidomics study for which we have proliferation rates
proliferation_metadata <- metadata %>%
  dplyr::select(group, Morphology, ProliferationRate) %>%
  distinct() %>%
  filter(group %in% data_cell_lines_notScaled$group) %>%
  filter(!is.na(ProliferationRate)) %>%
  arrange(group) %>%
  filter(Morphology %in% c("Epithelial-like", "Mesenchymal-like"))
dim(proliferation_metadata)

# Create a design object
design_morphology_prol <- model.matrix(
  ~ 0 + proliferation_metadata$Morphology + proliferation_metadata$ProliferationRate +
    proliferation_metadata$Morphology:proliferation_metadata$ProliferationRate
)

rownames(design_morphology_prol) <- proliferation_metadata$group
colnames(design_morphology_prol) <- str_replace_all(colnames(design_morphology_prol), "proliferation_metadata\\$", "")
colnames(design_morphology_prol) <- str_replace_all(colnames(design_morphology_prol), "Morphology", "")
colnames(design_morphology_prol) <- str_replace_all(colnames(design_morphology_prol), "-like", "")
colnames(design_morphology_prol) <- str_replace_all(colnames(design_morphology_prol), ":", ".") # needed for contrast fits
head(design_morphology_prol)

# Keep only cell lines for which we have proliferation rates
dataset <- data_cell_lines_notScaled %>%
  filter(group %in% proliferation_metadata$group) %>%
  column_to_rownames("group")
dim(dataset)
dataset[1:5, 1:5]
stopifnot(all(proliferation_metadata$group == rownames(dataset)))

# Fit the linear model using the LIMMA package
fit <- lmFit(t(dataset), design_morphology_prol)
fit <- eBayes(fit)

## Look at morphology effect
cm <- makeContrasts(
  Mesenchymal_proliferation = ProliferationRate + Mesenchymal.ProliferationRate,
  Epithelial_proliferation = ProliferationRate,
  Difference_ME = Mesenchymal - Epithelial,
  levels = design_morphology_prol
)
cm
fit_morphologies <- contrasts.fit(fit, cm)
fit_morphologies <- eBayes(fit_morphologies)

(all_contrasts <- colnames(cm))
names(all_contrasts) <- colnames(cm)

# Add metadata on lipid names and extract statistics for each contrast in the design
contrasts_lipids <- lapply(all_contrasts, function(x) {
  topTable(fit_morphologies, coef = x, n = Inf) %>%
    rownames_to_column("internal_identifiers") %>%
    left_join(lipid_info, by = "internal_identifiers") %>%
    arrange(P.Value)
})

head(contrasts_lipids$Mesenchymal_proliferation)

decide_direction <- function(logFC) {
  if (logFC < 0) {
    return("neg")
  }
  if (logFC > 0) {
    return("pos")
  }
  stop("something wrong")
}

decide_significance <- function(P.Value, adj.P.Val, logFC,
                                p_cutoff = 0.05, FDR_cutoff = 0.1) {
  direction <- decide_direction(logFC)
  if (adj.P.Val < FDR_cutoff) {
    return(str_c("sig ", direction))
  } else if (P.Value < p_cutoff) {
    return(str_c("almost sig ", direction))
  } else {
    return("not sig")
  }
  stop("Something wrong")
}

proliferation_lipids <- inner_join(
  dplyr::select(
    contrasts_lipids$Mesenchymal_proliferation,
    internal_identifiers, identifiers, subclass, logFC, P.Value, adj.P.Val
  ) %>%
    dplyr::rename(coefficient_prol = logFC),
  dplyr::select(
    contrasts_lipids$Epithelial_proliferation,
    internal_identifiers, identifiers, logFC, P.Value, adj.P.Val
  ) %>%
    dplyr::rename(coefficient_prol = logFC),
  by = c("internal_identifiers", "identifiers"), suffix = c("_M", "_E")
) %>%
  rowwise() %>%
  mutate(color_FDR = decide_significance(P.Value_E, adj.P.Val_E, coefficient_prol_E,
    p_cutoff = 0.05, FDR_cutoff = 0.1
  ))

head(proliferation_lipids)
```

#### Supplementary Table 4

```{r}
sheet1 <- metadata
sheet1 %>%
  write_csv(here(output_dir, "Supplementary_Table_4_sample_metadata.csv"))

sheet2 <- lipid_info %>%
  dplyr::select(-internal_identifiers) %>%
  relocate(class, subclass, .after = "total_chain_length")
sheet2 %>%
  write_csv(here(output_dir, "Supplementary_Table_4_lipid_metadata.csv"))

sheet3 <- data_cell_lines_notScaled %>%
  pivot_longer(!group, names_to = "internal_identifiers", values_to = "value") %>%
  pivot_wider(id_cols = internal_identifiers, names_from = group, values_from = value) %>%
  left_join(dplyr::select(lipid_info, internal_identifiers, identifiers, subclass, database_identifier),
    by = "internal_identifiers"
  ) %>%
  relocate(identifiers, subclass, database_identifier) %>%
  dplyr::select(-internal_identifiers)
sheet3 %>%
  write_csv(here(output_dir, "Supplementary_Table_4_normalized_lipids.csv"))

sheet4 <- proliferation_lipids %>%
  left_join(dplyr::select(lipid_info, internal_identifiers, database_identifier),
    by = "internal_identifiers"
  ) %>%
  relocate(identifiers, subclass) %>%
  dplyr::select(-internal_identifiers)
sheet4 %>%
  write_csv(here(output_dir, "Supplementary_Table_4_lipid_proliferation.csv"))

here(output_dir, "Supplementary_Table_4.xlsx")
writexl::write_xlsx(
  list(
    sample_metadata = sheet1, lipid_metadata = sheet2,
    normalized_lipids = sheet3, lipid_proliferation = sheet4
  ),
  path = here(output_dir, "Supplementary_Table_4.xlsx"),
  col_names = TRUE,
  format_headers = TRUE
)
```

#### Odds ratios and examples

Print statistics for the manuscript

```{r}
str_glue("Total nr of lipids: {nrow(proliferation_lipids)}")
str_glue("Total nr of lipids with epithelial FDR < 0.1: {nrow(filter(proliferation_lipids, adj.P.Val_E < 0.1))}")
str_glue("Total nr of lipids with mesenchymal FDR < 0.1: {nrow(filter(proliferation_lipids, adj.P.Val_M < 0.1))}")
```

Calculate odds ratios of side chains in the list of significantly associated lipids.
First, look at multiple properties and then check the properties that look of interest.

Check if a class as a whole is overrepresented in the list of significant hits.

Note that a low conditional odds ratio means "less than expected".

```{r}
proliferation_lipids_E <- contrasts_lipids$Epithelial_proliferation %>%
  mutate(direction = ifelse(logFC >= 0, "faster", "slower")) %>%
  mutate(subclass = factor(subclass))

proliferation_lipids_E %>%
  calculate_column_overrepresentations("subclass", direction = "faster", adj.P.Val = 0.1)

proliferation_lipids_E %>%
  calculate_column_overrepresentations("subclass", direction = "slower", adj.P.Val = 0.1)
```

Check for specific side chains

```{r}
# Subset of PUFA species and odd-chain fatty acids in slower-proliferating cells. And MUFA 20:1
for (pattern in c("15", "17", "20.3", "20.4", "20.5", "22.5", "22.6", "20.1")) {
  print(str_glue("hits for FA {pattern}"))
  contains_pattern <- count_pattern_in_limma_result(proliferation_lipids_E, pattern, "slower", adj.P.Val = 0.1)
  print(contains_pattern)
  contains_pattern <- calc_overrepresentation(contains_pattern)
  print(str_glue("FA {pattern} overrepresentation {round(contains_pattern$conditional_odds_ratio, 2)} ({round(contains_pattern$conditional_odds_ratio, 1)}) with p {round(contains_pattern$p_value_fisher, 3)}"))
}

# MUFA in faster-proliferating cells. And the previously mentioned PUFA and odd-chain FA.
for (pattern in c("15", "17", "20.3", "20.4", "20.5", "22.5", "22.6", "20.1")) {
  print(str_glue("hits for FA {pattern}"))
  contains_pattern <- count_pattern_in_limma_result(proliferation_lipids_E, pattern, "faster", adj.P.Val = 0.1)
  print(contains_pattern)
  contains_pattern <- calc_overrepresentation(contains_pattern)
  print(str_glue("FA {pattern} overrepresentation {round(contains_pattern$conditional_odds_ratio, 2)} ({round(contains_pattern$conditional_odds_ratio, 1)}) with p {round(contains_pattern$p_value_fisher, 3)}"))
}
```

Examples mentioned in the manuscript

```{r}
proliferation_lipids_E %>%
  filter(identifiers %in% c(
    "PI 18:1_20:1", "PI 18:0_20:1", "PI 16:0_20:1", "PI 18:0_18:1",
    "PI 18:1_18:1",
    "PI 16:1_18:0", "16:0_18:1",
    "PI 18:0_18:3", "PI 15:0_22:4", "PI 17:0_20:3", "PI 17:0_20:4",
    "PI18_0:20:3", "PI 18:0_20:4",
    "LPI 17:0", "LPI 20:4", "LPI 20:3"
  )) %>%
  dplyr::select(identifiers, logFC, adj.P.Val) %>%
  mutate(
    adj.P.Val = round(adj.P.Val, 3),
    logFC = round(logFC, 2)
  )
```

### Figure 5

#### Load data

For this manuscript we are using a dataset with and without treatment with a 2-Deoxy-D-glucose inhibitor. 
For this manuscript we only use control samples, with no drug added. This is why the dataset is called "treatment_2DG".

```{r}
data_treatment_2DG <- read_csv(here(data_dir, "data_treatment_2DG_log2Transformed_autoScaled.csv"))
data_treatment_2DG_t <- read_csv(here(data_dir, "data_treatment_2DG_t_log2Transformed_notScaled.csv")) %>%
  column_to_rownames("internal_identifiers")
metadata_timecourse <- read_csv(here(data_dir, "metadata_timecourse.csv"))
```

#### Figure 5A

Calculate lipid class averages. 
For each replicate and cell line, scale the changes of the class average to T0=0.

```{r}
lipid_class_averages_long <- data_treatment_2DG %>%
  calculate_lipid_class_average(lipid_info, subclass = T)

head(metadata_timecourse)
duplicated_metadata <- duplicate_control_T0_metadata(metadata_timecourse)

average_data_treatment_2DG <- lipid_class_averages_long %>%
  duplicate_control_T0_samples_long() %>%
  left_join(duplicated_metadata, by = "Original_name") %>%
  mutate(
    Subject = str_glue("{cell_line}_{biological_replicate}_{Treatment}"),
    sample = str_glue("{cell_line}_{biological_replicate}")
  ) ->
lipid_class_averages_over_time

lipid_class_averages_over_time %>%
  filter(Treatment == "ctrl") %>%
  # Calculate average for all cell lines and replicates
  group_by(cell_line, Treatment, Time, subclass) %>%
  summarise(
    subclass_sd = sd(subclass_mean),
    subclass_mean = mean(subclass_mean)
  ) %>%
  # Scale to set T0 as 0
  group_by(cell_line, subclass) %>%
  group_modify(scale_lipid_to_T0is0,
    lipid_column = "subclass_mean",
    amount_of_controls = 2
  ) %>%
  mutate(Subject = str_c(cell_line, Treatment)) -> lipid_class_averages_over_time

write_csv(lipid_class_averages_over_time, here(output_dir, "lipid_class_averages_scaledT0_over_time.csv"))
here(output_dir, "lipid_class_averages_over_time.csv")
```

#### Figure 5BCF

Calculate differences in lipid abundance over time. 

Part of the code for the design matrix is adapted from the tutorial of [RNAseq123](https://bioconductor.org/packages/release/workflows/vignettes/RNAseq123/inst/doc/designmatrices.html#studies-with-covariates-1) [DOI:10.18129/B9.bioc.RNAseq123](https://doi.org/doi:10.18129/B9.bioc.RNAseq123).

```{r}
## Create design formula
design_df <- metadata_timecourse %>%
  dplyr::select(Original_name, cell_line, Treatment, biological_replicate, Time, Subject) %>%
  mutate(Treatment_group = factor(str_c("t", Treatment, "_", Time, sep = ""),
    levels = c("tctrl_0", "tctrl_6", "tctrl_24", "tctrl_48", "t2DG_6", "t2DG_24", "t2DG_48")
  ))
head(design_df)
design_withcell <- model.matrix(~ 0 + design_df$Treatment_group + design_df$cell_line) %>%
  as.data.frame() %>%
  rename_with(function(x) str_remove(x, "design_df\\$Treatment_group")) %>%
  rename_with(function(x) str_replace(x, "tctrl", "ctrl")) %>%
  rename_with(function(x) str_remove(x, "design_df\\$cell_line")) %>%
  rename_with(function(x) str_remove_all(x, "-")) %>%
  as.matrix()

rownames(design_withcell) <- design_df$Original_name
colnames(design_withcell)

## Rename interaction term with a dot because the : gives errors later on
colnames(design_withcell) <- str_replace(colnames(design_withcell), ":", ".")

stopifnot(colnames(data_treatment_2DG_t) == design_df$Original_name)
corfit_withcell <- duplicateCorrelation(data_treatment_2DG_t, design_withcell, block = design_df$Subject)
corfit_withcell$consensus
fit_obj_with_cell_line <- lmFit(
  object = data_treatment_2DG_t, design = design_withcell,
  block = design_df$Subject, correlation = corfit_withcell$consensus
)
fit_obj_with_cell_line <- eBayes(fit_obj_with_cell_line)
colnames(fit_obj_with_cell_line)

## Look at treatment and time effects
cm <- makeContrasts(
  treatment_diff_6h = t2DG_6 - ctrl_6,
  treatment_diff_24h = t2DG_24 - ctrl_24,
  treatment_diff_48h = t2DG_48 - ctrl_48,
  time_diff_6_0 = ctrl_6 - ctrl_0,
  time_diff_24_0 = ctrl_24 - ctrl_0,
  time_diff_48_0 = ctrl_48 - ctrl_0,
  levels = design_withcell
)
cm
fit2 <- contrasts.fit(fit_obj_with_cell_line, cm)
fit2 <- eBayes(fit2)

lipid_48h_differences <- topTable(fit2, coef = "time_diff_48_0", n = Inf) %>%
  rownames_to_column("internal_identifiers") %>%
  left_join(
    dplyr::select(
      lipid_info, internal_identifiers, identifiers, subclass,
      total_chain_length, saturation, database_identifier
    ),
    by = "internal_identifiers"
  ) %>%
  mutate(subclass = ifelse(subclass == "unidentified sphingolipid", "unidentified", subclass)) %>%
  rowwise() %>%
  mutate(color_FDR = decide_significance(P.Value, adj.P.Val, logFC,
    p_cutoff = 0.05, FDR_cutoff = 0.1
  ))
dim(lipid_48h_differences)
head(lipid_48h_differences)
```

#### Supplementary Table 5

```{r}
sheet1 <- metadata_timecourse
sheet1 %>%
  write_csv(here(output_dir, "Supplementary_Table_5_sample_metadata.csv"))

sheet2 <- lipid_info %>%
  dplyr::select(-internal_identifiers) %>%
  relocate(class, subclass, .after = "total_chain_length")
sheet2 %>%
  write_csv(here(output_dir, "Supplementary_Table_5_lipid_metadata.csv"))

sheet3 <- data_treatment_2DG_t %>%
  rownames_to_column("internal_identifiers") %>%
  left_join(dplyr::select(lipid_info, internal_identifiers, identifiers, subclass, database_identifier),
    by = "internal_identifiers"
  ) %>%
  relocate(identifiers, subclass, database_identifier) %>%
  dplyr::select(-internal_identifiers)
sheet3 %>%
  write_csv(here(output_dir, "Supplementary_Table_5_normalized_lipids.csv"))

sheet4 <- lipid_48h_differences %>%
  dplyr::select(
    identifiers, subclass, total_chain_length, saturation,
    logFC, adj.P.Val, database_identifier, color_FDR
  )
sheet4 %>%
  write_csv(here(output_dir, "Supplementary_Table_5_lipid_time_differences_48h.csv"))

here(output_dir, "Supplementary_Table_5.xlsx")
writexl::write_xlsx(
  list(
    sample_metadata = sheet1, lipid_metadata = sheet2,
    normalized_lipids = sheet3, lipid_time_differences_48h = sheet4
  ),
  path = here(output_dir, "Supplementary_Table_5.xlsx"),
  col_names = TRUE,
  format_headers = TRUE
)
```

#### Figure 5D

```{r}
all_replicates_lipidtox <- read_csv(here(data_dir, "lipidtox_quantification_all_replicates.csv"))
# no normalization
all_replicates_mean_lipidtox <- all_replicates_lipidtox %>%
  group_by(cell_line, time, replicate) %>%
  summarize(
    mean_lipidtox = mean(Intensity_MeanIntensity_lipidtox),
    sd_lipidtox = sd(Intensity_MeanIntensity_lipidtox)
  ) %>%
  ungroup()

# Scale to T0 by adding the matching T0 sample for each replicate.
all_replicates_mean_lipidtox_scaled <- left_join(
  all_replicates_mean_lipidtox,
  all_replicates_mean_lipidtox %>%
    filter(time == 0) %>%
    dplyr::select(cell_line, replicate, mean_lipidtox) %>%
    dplyr::rename(mean_lipidtox_T0 = mean_lipidtox),
  by = c("cell_line", "replicate")
) %>%
  mutate(scaled_lipidtox = mean_lipidtox / mean_lipidtox_T0)

# Average per cell line
scaled_lipidtox_averages <- all_replicates_mean_lipidtox_scaled %>%
  group_by(cell_line, time) %>%
  summarize(
    mean_scaled_lipidtox = mean(scaled_lipidtox),
    sd_lipidtox = sd(scaled_lipidtox)
  ) %>%
  ungroup() %>%
  dplyr::rename(Time = time) %>%
  mutate(log2FC = log2(mean_scaled_lipidtox))
head(scaled_lipidtox_averages)

# Save. This dataset contains averaged lipidtox intensities for three cell lines.
scaled_lipidtox_averages %>%
  write_csv(here(output_dir, "cell_line_averaged_lipidtox.csv"))
```
#### Figure 5E

```{r}
file.copy(
  here(data_dir, "lipidtox_image.tiff"),
  here(output_dir, "lipidtox_image.tiff")
)
```

## Supplementary Figures

### Supplementary Figure 1

```{r}
file.copy(
  here(data_dir, "unknown_sphingolipid_example.xlsx"),
  output_dir
)
```

### Supplementary Figure 2

Data is created together with data for Figure 3.

### Supplementary Figure 3

```{r}
calculate_SFA_MUFA_PUFA <- function(df_cell_lines, type_calculation) {
  data_cell_lines_t <- pivot_longer(df_cell_lines, !group, names_to = "internal_identifiers") %>%
    pivot_wider(id_cols = internal_identifiers, names_from = "group", values_from = "value")

  print(dim(data_cell_lines_t))
  lipid_info %>%
    dplyr::select(internal_identifiers, subclass, saturation) %>%
    left_join(data_cell_lines_t, by = "internal_identifiers") %>%
    dplyr::select(-internal_identifiers) %>%
    group_by(subclass, saturation) -> df_cell_lines_saturation
  if (type_calculation == "sum") {
    df_cell_lines_saturation %>%
      summarize(across(everything(), sum)) %>%
      ungroup() -> df_cell_lines_saturation
  } else if (type_calculation == "mean") {
    df_cell_lines_saturation %>%
      summarize(across(everything(), mean)) %>%
      ungroup() -> df_cell_lines_saturation
  } else if (type_calculation == "percentage") {
    totals <- df_cell_lines_saturation %>%
      ungroup() %>%
      group_by(subclass) %>%
      dplyr::select(-saturation) %>%
      summarize(across(everything(), sum)) %>%
      ungroup()
    saturations <- df_cell_lines_saturation %>%
      summarize(across(everything(), sum)) %>%
      ungroup()
    df_cell_lines_saturation <- left_join(
      pivot_longer(saturations, !c(subclass, saturation), names_to = "cell_line", values_to = "saturation_sum"),
      pivot_longer(totals, !subclass, names_to = "cell_line", values_to = "totals_sum"),
      by = c("subclass", "cell_line")
    ) %>%
      mutate(percentage = saturation_sum / totals_sum)
  }
  print(dim(df_cell_lines_saturation))
  return(df_cell_lines_saturation)
}
```

```{r}
raw_cell_line_data <- read_csv(here(data_dir, "cell_lines_replicateAveraged_PQNNormalized_notTransformed_notScaled.csv"))
raw_cell_line_data_saturation <- calculate_SFA_MUFA_PUFA(raw_cell_line_data, type_calculation = "percentage")

raw_cell_line_data_saturation %>%
  left_join(metadata, by = c("cell_line" = "group")) %>%
  arrange(Morphology) %>%
  mutate(cell_line = factor(cell_line, levels = pull(., cell_line) %>% unique())) %>%
  filter(subclass %in% c("PE", "PS", "PI", "PC")) -> saturation_percentages

write_csv(saturation_percentages, here(output_dir, "saturation_percentages.csv"))
```

## Additional information

### GO database version

```{r}
org.Hs.eg.db
```
### Citation of versions

```{r}
knitr::write_bib(c(
  .packages(),
  "mixOmics"
), here(output_dir, "packages_calculations.bib"))
```

### Session info

```{r}
sessionInfo()
```
